Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2021-11-16 src/ms.tex
ok now I think we actually have a working draft
updated workflow

% Abstract with filler text
\begin{abstract}
   Short-period double white dwarf (DWD) binaries will be the most prolific source of gravitational waves (GWs) for the Laser Interferometer Space Antenna (LISA). Not only will tens of thousands of DWDs be individually resolved, but DWDs with GW frequencies below $\sim1\,\rm{mHz}$ will be the dominant contributor to a stochastic foreground caused by confusion from overlapping GW signals, limiting the detectability of individual sources of all kinds. The number of DWDs in the Milky Way strongly depends on the Galactic stellar binary fraction. Population modelling of Galactic DWDs typically assumes a standard binary fraction of $50\%$. However, recent observations have shown that the binary fraction of close ($P_{\text{orb}}$ $\leq 10^4$ days) solar-type stars has a strong anti-correlation with stellar metallicity. In this study we perform the first simulation of the Galactic DWD population observable by LISA which incorporates an empirically-derived metallicity-dependent binary fraction. We simulate DWDs using the binary population synthesis suite \cosmic\ and scale our simulations to a Galactic population using a metallicity-dependent star formation history to create populations of DWDs which orbit in the LISA band. We find that while metallicity impacts the evolution and intrinsic properties of our simulated DWD progenitor binaries, the properties of the Galactic population of close DWDs remain unchanged when considering a metallicity-dependent binary fraction. However, the size of the Galactic DWD population is reduced by more than half when accounting for a metallicity-dependent binary fraction. This effect only slightly decreases the number of DWDs individually resolved by LISA and also serves to lower the confusion foreground, effectively increasing the sensitivity for detecting low-frequency LISA sources of all kinds. We repeat our analysis for three different assumptions for Roche-lobe overflow interactions and find the population reduction to be robust when a metallicity-dependent binary fraction is assumed.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the stars which are born with a binary companion, many will undergo interactions which bring the two stars closer together, eventually forming a close double white dwarf (DWD). Close DWDs, with orbital periods shorter than $\sim2.75\,\rm{hr}$, are the largest source by number of mHz gravitational waves (GWs) in the Galaxy \citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space Antenna (LISA) is expected to resolve at least $10^4$ individual DWD binaries in the Milky Way and will also observe GW emission from the entire Galactic DWD population through the unresolved foreground created by overlapping signals at sub-mHz frequencies \citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, Korol2017, Lamberts2019, Breivik2020a}. The resolved population will enable the study of several important aspects of binary evolution like the strength of tides \citep{Valsecchi2012} and the stability of mass transfer in DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, Sepinsky2014, Kremer2015} as well as provide a probe of Galactic structure \citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and strength of the Galactic DWD foreground can also be used as a tool to study the structure of the Milky Way \citep{Benacquista2006, Breivik2020b} as well as the separation distribution of close DWDs \citep{Korol2021}.
When viewed strictly as a source of noise, the unresolved Galactic DWD foreground is the dominant noise source for LISA in the sub-mHz part of LISA's frequency range. This extra noise above the detector noise affects the detection of all other LISA sources including extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, merging black holes with masses between $10^4\,\rm{M_{\odot}}$--$10^7\,\rm{M_{\odot}}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. For sources which have signals buried by the Galactic DWD population, the foreground must be carefully analyzed and subtracted \citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number of resolved DWDs and the height of the unresolved DWD foreground are a direct consequence of the number of DWD progenitors which form and evolve over the Milky Way's history.

While the binary fraction remains approximately constant across a large metallicity range ($-1.5 \leq$ [Fe/H] $< 0.5$) for wide binaries, close OB stars, and the stellar Initial Mass Function (IMF) \citep{Moe2017, Moe2019}, the binary fraction for solar-type star systems with orbital period $P_{\text{orb}}\leq 10^4$ days (separation $a \leq 10$ AU) shows a strong anti-correlation with metallicity \citep[e.g.][]{Badenes2018, Moe2019, Mazzola2020, Price-Whelan2020}. Because close DWDs are the remnants of close, solar-type, binary stars, this anticorrelation plays an important role in the formation, evolution, and characteristics of the DWD population that LISA will observe.

To date, population synthesis studies of the Galactic population of close DWDs have either assumed a $100\%$ binary fraction or a $50\%$ binary fraction, such that for every three stars formed, two reside in a binary system \citep{Nelemans2001a, Yu2013, Korol2017, Lamberts2019}. In this study, we investigate the effects of a metallicity-dependent binary fraction on the formation and evolution of DWDs. To this end, we create synthetic present day Milky Way-like galaxies of DWDs and specifically select systems with GW signals that may be observable by the space-based detector LISA. Throughout, we make comparisons between the standard assumption of a constant $50\%$ initial binary fraction (hereafter model F50) and one with a metallicity-dependent binary fraction (hereafter model FZ).

In Section \ref{sec:simulations} we discuss our binary evolution assumptions used in simulating DWD populations and detail the process to produce present day synthetic Milky-Way-like galaxies. In Section \ref{sec:LISA_obs} we review the derivation of LISA detectability for circular DWD populations at mHz frequencies. In Section \ref{sec:results} we detail results showing how a metallicity-dependent binary fraction affects the formation and evolution of DWD populations assuming a fiducial set of binary evolution assumptions. In Section \ref{sec:LISA_met} we detail the metallicity dependence of the LISA DWD foreground and resolved population. Finally, we repeat our analysis for three sets of binary evolution assumptions to show a robust reduction of the size of Galactic DWD foreground in LISA for populations simulated with a metallicity dependent binary fraction in Section \ref{sec:model_compare} and conclude in Section \ref{sec:conclusions}.

@@

\subsection{Binary population models}
\label{sec:bin_pop}
We simulate the evolution of DWD progenitor populations using \cosmic\footnote{https://cosmic-popsynth.github.io}, an open-source Python-based rapid binary population synthesis suite which employs single and binary star evolution using \texttt{SSE/BSE} \citep{Hurley2000, Hurley2002}. Several modifications have been added to \cosmic\ which incorporate updates for massive star evolution and binary interactions. For a detailed description of these modifications see \citet{Breivik2020a}. \cosmic\ has been used in several studies to study the effects of binary evolution on binary populations from blue stragglers \citep{Leiner2021} and heartbeat stars \citep{Jayasinghe2021}, to white dwarf populations \citep{Kremer2017,Breivik2018,Kilic2021}, to merging compact object populations in isolated binaries \citep{Zevin2020b, Zevin2020a, Zevin2021, Wong2021, Mandhai2021} and in dynamical environments around super-massive black holes \citep{Stephan2019, Wang2021}.

@@

@@

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE models}
\label{sec:FIRE}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/SFH_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close solar-type binaries with P $< 10^4$ days (black) plotted against the logarithm of Z/Z$_\odot$, with scatter points denoting the location of the metallicity grid used in our \cosmic\ simulations. The secondary axis shows the amount of mass in stars formed within each metallicity bin from galaxy \textbf{m12i} of the ``Latte" suite in FIRE as a red histogram. The amount of stellar mass formed with super-solar metallicities dominates the distribution. Note that the primary axis shows $f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of mass formed in log-scale. The opposing trends of these two distributions compete throughout this study.}
    \label{fig:SFH_vs_fb}
\end{figure}

The metallicity-dependent binary fraction is shown in black in Figure \ref{fig:SFH_vs_fb} along with the mass in star particles from galaxy \textbf{m12i}, shown in red, as a function of our metallicity grid. The binary fraction, $f_{\rm{b}}$, drops drastically across metallicity while the mass formed in FIRE increases significantly. These two opposing trends compete within this study along with the impact of metallicity on single star evolution to form the final numerical distribution of systems in our DWD populations.

Since our \cosmic\ simulations assume a binary fraction of $f_{\rm{b}} = 1$, we scale the amount of mass sampled at ZAMS required to generate the population of DWDs to the proper binary fraction for each model. We do this by sampling single stars and primary masses of binary stars from the \cite{Kroupa2001} IMF and sampling secondary masses of the binary stars from a uniform mass distribution, where the number of binaries is determined by the binary fraction model (F50 or FZ). From this sample, we obtain the ratio of mass in single stars to the mass in binary stars as a function of the binary fraction, $R(f_{\rm{b}})$. For model F50, the ratio is a constant $R(f_{\rm{b}}) =$ 0.64. For model FZ, the ratio increases from 0.68 at low metallicities to 3.17 at solar metallicity indicating fewer mass in binaries relative to single stars. This ratio can then be used to scale the total amount of ZAMS mass sampled in binary stars required to produce our simulated population from \cosmic\ to $M_{\rm{ZAMS,sim}}$, the total amount of ZAMS mass in single stars and binary systems required to produce the simulated population for a given binary fraction $f_{\rm{b}}$.

@@

\noindent The number of DWD’s per FIRE star particle at metallicity $Z_i$ is then
\begin{equation}
    N_{\rm{DWD},\star} (Z_i) = n_{\rm{DWD}}(Z_i)\, M_\star,
\end{equation}

\noindent where $M_\star$ = 7070M$_\odot$ is the mass per FIRE star particle for the \textbf{m12i} galaxy. Since $N_{\rm{DWD},\star} (Z_i)$ is not an integer, we treat the decimal component as the probability that the star particle contains an extra DWD beyond the integer number. For each star particle, we sample with replacement $N_{\rm{DWD},\star}(Z_i)$ DWDs from the corresponding simulated \cosmic\ population at that metallicity and assign the ZAMS birth time of each DWD to the formation time of the star particle. For most DWD types there is more than one DWD binary system assigned to each \textbf{m12i} star particle.

If the DWD formation time is less than the age of the star particle, we evolve the DWD over the remaining time between its formation and star particle age, $t_{\rm{evol}}$, to produce the present day population. Once a DWD is formed, we assume that the binary evolves only due to the emission of GWs. Due to tidal effects and mass transfer between their progenitor binaries, all DWDs in our simulations are circular, thus eccentricity does not need to be considered. The orbital evolution over the time $t_{\rm{evol}}$ is then simply defined according to \citet{Peters1964} as
\begin{equation}
    a_f = (a_i - 4\beta t_\text{evol})^{1/4},
\end{equation}
where where $a_i$ is the DWD separation at formation, and 
\begin{equation}
    \beta = \frac{64G^3}{5c^5} M_1M_2(M_1+M_2)
\end{equation}
is constant throughout DWD evolution \citep{Peters1964}.

@@

\begin{equation}
    a_{\text{RLO,}\ell} = R_{\ell} \cdot \frac{0.6 q_{\ell}^{2/3} + \ln{(1+q_{\ell}^{1/3})}}{0.49 q_{\ell}^{2/3}}
\end{equation}
where $R_{\ell}$ is the radius of the lower mass WD and $q_{\ell} = M_{\ell}/M_{h}$ is the ratio of the lower- to higher-mass WD components. \citep{Eggleton1983}. 
We define the radius of a WD following \citet{Tout1997, Hurley2000} as
\begin{equation}
    R = \max\Bigg(R_\text{NS}, 0.0115\sqrt{\left(\frac{M_\text{Ch}}{M}\right)^{2/3}-\left(\frac{M}{M_\text{Ch}}\right)^{2/3}} \Bigg) 
\end{equation}
\noindent where $R_\text{NS} = 1.4\cdot 10^{-5}$ R$_\odot$ is the radius of a neutron star, $M_{\text{Ch}}=1.44$ M$_\odot$ is the Chandrasekhar limit for the mass of a stable WD, and $M$ is the mass of the DWD in solar masses.

For the non-discarded systems, we log the present day separations from which the present day orbital frequency $f_\text{orb}$ can be found using Kepler's third law. The GW frequency is then $f_\text{GW} = 2f_\text{orb}$.

>>>>

% Abstract with filler text
\begin{abstract}
   Short-period double white dwarf (DWD) binaries will be the most prolific source of gravitational waves (GWs) for the Laser Interferometer Space Antenna (LISA). Not only will tens of thousands of DWDs be individually resolved, but DWDs with GW frequencies below $\sim1\,\rm{mHz}$ will be the dominant contributor to a stochastic foreground caused by confusion from overlapping GW signals, limiting the detectability of individual sources of all kinds. The number of DWDs in the Milky Way strongly depends on the Galactic stellar binary fraction. Population modelling of Galactic DWDs typically assumes a standard binary fraction of $50\%$. However, recent observations have shown that the binary fraction of close ($P_{\text{orb}}$ $\leq 10^4$ days) solar-type stars exhibits a strong anti-correlation with metallicity. In this study we perform the first simulation of the Galactic DWD population observable by LISA which incorporates an empirically-derived metallicity-dependent binary fraction. We simulate DWDs using the binary population synthesis suite \cosmic\ and incorporate a metallicity-dependent star formation history to create a Galactic population of short-period DWDs. We compare two models: one which assumes a metallicity-dependent binary fraction, and one with a binary fraction of 50\%. We find that while metallicity impacts the evolution and intrinsic properties of our simulated DWD progenitor binaries, the LISA-resolvable populations of the two models remain roughly indistinguishable. However, the size of the total Galactic DWD population orbiting in the LISA frequency band is reduced by more than half when accounting for a metallicity-dependent binary fraction. This effect serves to lower the confusion foreground, effectively increasing the sensitivity for detecting all types of low-frequency LISA sources. We repeat our analysis for three different assumptions for Roche-lobe overflow interactions and find the population reduction to be robust when a metallicity-dependent binary fraction is assumed.

@@

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the stars which are born with a binary companion, many will undergo interactions which bring the two stars closer together, eventually forming a close double white dwarf (DWD). Close DWDs, with orbital periods shorter than $\sim2.75\,\rm{hr}$ %\snotes{[close DWDs are defined to have $P_{\rm{orb}}\leq10^4$ days - are we noting 2.75 hrs for GW sources specifically? If so maybe we should reword this bit because it reads like it conflicts with the abstract?]},
%Katie: Close DWDs is a different classification than close binaries.
are the largest source by number of mHz gravitational waves (GWs) in the Galaxy \citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space Antenna (LISA) is expected to resolve at least $10^4$ individual DWD binaries in the Milky Way and will also observe GW emission from the entire Galactic DWD population through the unresolved foreground created by overlapping signals at sub-mHz frequencies \citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, Korol2017, Lamberts2019, Breivik2020a}. The resolved population will enable the study of several important aspects of binary evolution like the strength of tides \citep{Valsecchi2012} and the stability of mass transfer in DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, Sepinsky2014, Kremer2015} as well as provide a probe of Galactic structure \citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and strength of the Galactic DWD foreground can also be used as a tool to study the structure of the Milky Way \citep{Benacquista2006, Breivik2020b} as well as the separation distribution of close DWDs \citep{Korol2021}.

When viewed strictly as a source of noise, the unresolved Galactic DWD foreground is the dominant noise source for LISA in the sub-mHz part of LISA's frequency range. This extra noise above the detector noise floor affects the detection of all other LISA sources including extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, merging black holes with masses between $10^4$--$10^7\,M_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. For sources which have signals buried by the Galactic DWD population, the foreground must be carefully analyzed and subtracted \citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number of resolved DWDs and the height of the unresolved DWD foreground are a direct consequence of the number of DWD progenitors which form and evolve over the Milky Way's history.

While the binary fraction remains approximately constant across a large metallicity range ($-1.5 \leq$ [Fe/H] $< 0.5$) for wide binaries, close OB stars, and the stellar Initial Mass Function (IMF) \citep{Moe2017, Moe2019}, the binary fraction for solar-type star systems with orbital period $P_{\text{orb}}\leq 10^4$ days (separation $a \leq 10$ AU) shows a strong anti-correlation with metallicity \citep[e.g.][]{Badenes2018, Moe2019, Mazzola2020, Price-Whelan2020}. Because close DWDs are the remnants of close, solar-type binary stars, this anti-correlation plays an important role in the formation, evolution, and characteristics of the DWD population that LISA will observe.

%These relatively close binaries are likely to form from fragmentation, accretion, and inward migration in protostellar disks. This is in contrast to wide binaries whose formation channels stem from fragmentation of molecular cores, which is generally unaffected by the opacity of the material. 
%This is depicted in the ``metallicity invariance" for the IMF and binary fraction beyond separations of 200 au for -1.0 $<$ [Fe/H] $<$ 0.5 of Mazzola et al. (2020). Within protostellar disks however, it has been shown through analytical models and hydrodynamic simulations that the probability of disk fragmentation decreases with increasing metallicity (Mazzola et al. 2020). Metal-poor cores are unable to radiate through molecular transitions as effectively, and so are generally hotter and must reach higher masses in order to collapse into disks, which instigates higher accretion rates and results in gravitational instability and thus fragmentation. Furthermore, if you decrease a disk’s metallicity and thus its optical depth, the mid-plane of the disk is able to radiate and cool more efficiently which encourages fragmentation. (Mazzola et al. 2020 and others).

%Numerically, the dominant sources for LISA will be DWDs from the Milky Way. There is an expectation that several tens of millions of detached DWDs will be present in the LISA band with about ten thousand of these to be individually resolvable  with $f_\text{GW} \geq 0.4$ mHz (Nelemans et al. 2001b). DWDs are nice systems to work with: a compact binary which emits continuously and near monochromatically GW signals in the source frame (LISA proposal REF), and of which the second harmonic of their signals can be measured with high accuracy. In general, higher frequency systems are louder and better characterized than low frequency counterparts. In fact, at low frequencies we find that the high number of DWDs begin to overlap in their signals, creating a stochastic background. This effective “confusion signal” arises for these low frequencies from having more than one DWD per frequency bin, $f_\text{bin} = 1 / t_\text{obs}$, where $t_\text{obs}$ is the length of observation time for the LISA mission, which we assume to be four years (LISA proposal REF).

%The science that the LISA mission will contribute to is diverse in scope. Specific to DWDs, the hope is to use the distribution of chirp masses and periods to constrain the impact of the common envelope, which drastically tightens the orbit of the systems (Toonen et al. 2014). A complete sample will also allow for direct comparison with the post-common envelope binaries, which have undergone only one episode of the common envelope (Rebassa-Mansergas et al. 2012). In some cases, the frequency derivative will be measurable and allow us to determine if mass transfer is happening (Breivik et al. 2018) and/or tidal interactions are deforming the white dwarfs.

%\subsection{The metallicity-dependent binary fraction}
To date, population synthesis studies of the Galactic population of close DWDs have either assumed a $100\%$ binary fraction or a $50\%$ binary fraction, such that for every three stars formed, two reside in a binary system \citep{Nelemans2001a, Yu2013, Korol2017, Lamberts2019}. In this study, we investigate the effects of a metallicity-dependent binary fraction on the formation and evolution of DWDs. To this end, we create synthetic present-day Milky Way-like galaxies of DWDs and specifically select systems with GW signals that may be observable by the space-based detector LISA. Throughout, we make comparisons between the standard assumption of a constant $50\%$ initial binary fraction (hereafter model F50) and one with a metallicity-dependent binary fraction (hereafter model FZ).

In Section \ref{sec:simulations} we discuss our binary evolution assumptions used in simulating DWD populations and detail the process to produce present-day synthetic Milky-Way-like galaxies. In Section \ref{sec:LISA_obs} we review the derivation of LISA detectability for circular DWD populations at mHz frequencies. In Section \ref{sec:results} we detail results showing how a metallicity-dependent binary fraction affects the formation and evolution of DWD populations assuming a fiducial set of binary evolution assumptions. In Section \ref{sec:LISA_met} we detail the metallicity dependence of the LISA DWD foreground and resolved population. Finally, we repeat our analysis for three sets of binary evolution assumptions to show a robust reduction of the size of the Galactic DWD foreground in LISA for populations simulated with a metallicity-dependent binary fraction in Section \ref{sec:model_compare} and conclude in Section \ref{sec:conclusions}.

@@

%The Latte suite of FIRE-2 cosmological zoom-in baryonic simulations of Milky Way-mass galaxies \citep{Wetzel2016}, part of the Feedback In Realistic Environments (FIRE) simulation project, were run using the Gizmo gravity plus hydrodynamics code in meshless finite-mass (MFM) mode \citep{Hopkins2015} and the FIRE-2 physics model \citep{Hopkins2018}. Synthetic Gaia DR2-like surveys of the Latte suite of FIRE-2 simulations were created via the Ananke framework \citep{Sanderson2020}.

\subsection{Binary population models}
\label{sec:bin_pop}
We simulate the evolution of DWD progenitor populations using \cosmic\footnote{https://cosmic-popsynth.github.io}, an open-source Python-based rapid binary population synthesis suite which employs single and binary star evolution using \texttt{SSE/BSE} \citep{Hurley2000, Hurley2002}. Several modifications have been added to \cosmic\ which incorporate updates for massive star evolution and binary interactions. For a detailed description of these modifications see \citet{Breivik2020a}. \cosmic\ has been used in several studies to examine the effects of binary evolution on binary populations from blue stragglers \citep{Leiner2021} and heartbeat stars \citep{Jayasinghe2021}, to white dwarf populations \citep{Kremer2017,Breivik2018,Kilic2021}, to merging compact object populations in isolated binaries \citep{Zevin2020b, Zevin2020a, Zevin2021, Wong2021, Mandhai2021} and in dynamical environments around super-massive black holes \citep{Stephan2019, Wang2021}. 
%Recently, \cosmic\ was integrated with the Cluster Monte Carlo (CMC) code to enable direct comparisons between isolated binary evolution and the evolution of stars in globular clusters \citep{Rodriguez2021}. Since this work considers the effects of the metallicity-dependence of the close isolated binary fraction, we leave direct comparisons to globular clusters to future work.

@@

% I'm suggesting text above since we have a boiler-plate style that we use to stay consisten in the literature.
%Rather than evolving an initial binary population’s individual stars - a computationally expensive process - \cosmic\ uses Monte Carlo statistics to create fixed populations that would be expected based on parameters like SFH and the known processes within binary star evolution. In the \cosmic\ framework, binaries are initialized iteratively with zero age main sequence (ZAMS) binary parameters and tested against the final parameter distributions desired by the user until they match. The output of \cosmic\ contains enough information that a user can analyze both individual systems of interest as well as the overall population, from ZAMS until compact binary formation.

%In this study, we used \cosmic\ 's cosmic-pop package to create fixed populations of DWDs, keeping all parameters except DWD type and metallicity Z constant. The output fully describes the binary parameter distributions that result from our SFH and binary evolution model, coined BSEDict, which is specified in an inifile containing sections on filters, convergence, sampling, and BSE flags, which will be described below:

%\katie{Based on the restructuring I did, I think you should expand this paragraph to describe (1) since we are interested in a metallicity-dependent binary fraction, we simulate metallicity bins separately, (2) since the masses and orbital periods of each DWD type at formation span a wide range, we simulate them separately.}
%For a given metallicity, a fixed population of one million binaries was created for each DWD subtype (discussed in Section \ref{sec:ini}). 
%\katie{Is this true? I think we probably simulated more. We should talk about this today!}

%moved paragraph up here from below since it takes care of some of the things I suggested in the paragraph above!

@@

%this is covered in my text above.
%We iteratively simulate binary populations until the distributions of the DWD masses and orbital period at the formation of the second-formed WD converge to a stable shape. Eccentricity does not need to be analyzed, as most DWDs are circularized by tidal forces (Marsh et al. 2004, Gokhale et al. 2007, Kremer et al. 2015/17), which also simplifies many calculations later on for GW evolution and GW signals.

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE models}
\label{sec:FIRE}
%\katie{I think this paragraph gives a bit too much detail on FIRE. I think the only detail we need is (1) we use Galaxy m12i from the Latte Suite of the FIRE simulations and (2) Latte reproduces parts of the Milky Way (see Lamberts+2019 paper for this info)}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations. %\sout{Specifically, the new “FIRE-2” code version was used, which updates the FIRE physics modules in the code GIZMO (Hopkins et al. 2018;  Gabyurov and Nitadori 2011; Hopkins 2015). This update leverages recent developments of mesh-free Godunov hydrodynamics methods, and incorporates updates to “cooling and recombination rates, gravitational force softening, and numerical feedback coupling” (Hopkins et al. 2018). The Latte suite is then run using Gizmo gravity, plus hydrodynamics code in meshless finite-mass (MFM) mode (Hopkins 2015) and the FIRE-2 physics model (Hopkins e al 2018)” $<$- citation}.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close solar-type binaries with P $< 10^4$ days (black) plotted against the logarithm of Z/Z$_\odot$, with scatter points denoting the location of the metallicity grid used in our \cosmic\ simulations. The secondary axis shows the amount of mass in stars formed within each metallicity bin from galaxy \textbf{m12i} of the ``Latte" suite in FIRE-2 simulations as a red histogram. The amount of stellar mass formed with super-solar metallicities dominates the distribution. Note that the primary axis shows $f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of mass formed in log-scale. The opposing trends of these two distributions compete throughout this study.}
    \label{fig:sfh_vs_fb}
\end{figure}

The metallicity-dependent binary fraction is shown in black in Figure \ref{fig:sfh_vs_fb} along with the mass in star particles from galaxy \textbf{m12i}, shown in red, as a function of our metallicity grid. %Plotted in black is the binary fraction as a function of metallicity. 
%The metallicity values chosen for our grid are shown as the scatter points. The red histogram depicts the total mass in stars formed for each metallicity bin within FIRE galaxy \textbf{m12i}, the Milky Way-like simulation used in this study which will be described further in Section \ref{sec:simulations}, as a function of metallicity. 
The binary fraction, $f_{\rm{b}}$, drops drastically across metallicity while the mass formed in \textbf{m12i} increases significantly. These two opposing trends compete throughout this study along with the impact of metallicity on single star evolution to form the final numerical distribution of systems in our DWD populations.

Since our \cosmic\ simulations assume a binary fraction of $f_{\rm{b}} = 1$, we scale the amount of mass sampled at ZAMS required to produce our \cosmic-generated population of DWDs ($M_{\text{b,ZAMS}}$) to the proper amount of mass sampled in single and binary stars ($M_{\text{ZAMS,sim}}$) for each binary fraction and model. We do this by sampling single stars and primary masses of binary stars from the \cite{Kroupa2001} IMF and sampling secondary masses of the binary stars from a uniform mass distribution, where the number of binaries is determined by the binary fraction model (F50 or FZ). From this sample, we obtain the ratio of mass in single stars to the mass in binary stars as a function of the binary fraction, $R(f_{\rm{b}})$. For model F50, the ratio is a constant $R(f_{\rm{b}}) = 0.64$. For model FZ, the ratio increases from 0.68 at low metallicities to 3.17 at solar metallicity indicating less mass in binaries relative to single stars. The total amount of ZAMS mass in single and binary stars is then $M_{\text{ZAMS,sim}} =  M_{\text{b,ZAMS}} (1 + R(f_{\rm{b}}))$.

%We use these ratios to then determine the total amount of ZAMS mass sampled in single and binary stars. 
%This ratio can then be used to scale the total amount of ZAMS mass sampled in binary stars required to produce our simulated population from \cosmic\ to $M_{\rm{ZAMS,sim}}$, the total amount of ZAMS mass in single stars and binary systems required to produce the simulated population for a given binary fraction $f_{\rm{b}}$.

%For a given metallicity in our grid and a given DWD type of one compact object binary type, the general procedure to create the Milky Way like galaxy is as follows. We specify metallicity bin $i$ with metallicities Z $\in$ [Z$_{i-1}$, Z$_i$]. The number of DWDs per unit mass N$_{\text{DWD,}i}$ for the \katie{\sout{cosmic-pop} simulated \cosmic} population pop$_i$ with Z = Z$_i$ is found through:

%\begin{align*}
%    R_{m}(f_i) &= \frac{M_{\text{singles,}\odot}(f_i)}{M_\text{binaries,} \odot(f_i)} \numberthis \\
%    \\
%    M_{\text{tot,}i} &=  M_{\text{binaries,}i} \cdot (1 + R_m(f_i)) \numberthis \\
%    \\
%    N_{\text{DWD,}i} &= \frac{N_{\text{DWD, tot,}i}}{M_{\text{tot,}i}} \numberthis
%\end{align*}

%R$_m$ is the fraction of mass in single stars to mass in binary stars for binary fraction $f_i$. \katie{I wonder if we can rewrite these equations to be a bit less confusing; I think we should use Lamberts+2019 as a guide.} \katie{\sout{This is calculated ahead of time using the \cosmic\ independent sampler in the InitialBinaryTable class.} I think that this sentence should say that the scaling incorporates the same assumptions for primary and secondary mass distributions as are used in the simulated data.} One million DWD binaries are initialized at solar metallicity for a given binary fraction, along with the total mass in singles vs. binaries. M$_{{\rm{binaries, }}i}$ the mass in binaries of pop$_i$ (the total mass of the population, as they were run with f$_{\rm{b}}$ = 1.0). M$_{{\rm{binaries, }}i}$ is scaled using R$_m$ to find the total astronomical mass M$_{\rm{tot, Z}_i}$ of the population. N$_{{\rm{DWD, tot,}}i}$ is the total number of DWDs in pop$_i$.

@@

\noindent The number of DWD’s per \textbf{m12i} star particle at metallicity $Z_i$ is then
\begin{equation}
    N_{\rm{DWD},\star} (Z_i) = n_{\rm{DWD}}(Z_i)\, M_\star,
\end{equation}

\noindent where $M_\star = 7070\,M_\odot$ is the mass per \textbf{m12i} star particle. Since $N_{\rm{DWD},\star} (Z_i)$ is not an integer, we treat the decimal component as the probability that the star particle contains an extra DWD in addition to the integer number. For each star particle, we sample with replacement $N_{\rm{DWD},\star}(Z_i)$ DWDs from the corresponding simulated \cosmic\ population at that metallicity and assign the ZAMS birth time of each DWD to the formation time of the star particle. For most DWD types there is more than one DWD binary system assigned to each \textbf{m12i} star particle.

%Already covered above
%In order to ensure that each DWD position is unique, we assign positions inside the star particle following the Ananke framework which provides offsets from the star particle center.

%\katie{I think this should go later}This gives us a final distance $d$ to the sun for all the systems, and from this and other information, we can calculate parameters relevant to the LISA observatory like Amplitude Spectral Density (ASD), signal-to-noise ratio (SNR), chirps, chirp masses, etc, as discussed in the next section.

%This will be a floating point number. The integer component gives one number of DWDs to sample with replacement from the cosmic-pop output to apply to each FIRE star particle. The decimal component will be treated as a probability $p$ of a binary existing for that particle. We assign an additional DWD to each star particle with probability $p$. The DWDs now have metallicities, ages, and galactic positions.

%\katie{I think the presentation of the next couple of paragraphs is a bit out of order. What about presenting it in the order: (1) we evolve the DWDs from their formation time up to the present day using the Peters equations (2) we discard any systems which overflow their roche lobes.}

If the DWD formation time is less than the age of the star particle, we evolve the DWD over the remaining time between its formation and star particle age, $t_{\rm{evol}}$, to produce the present-day population. Once a DWD is formed, we assume that the binary evolves only due to the emission of GWs. Due to tidal effects and mass transfer between their progenitor binaries, all DWDs in our simulations are circular, thus eccentricity does not need to be considered. The orbital evolution over the time $t_{\rm{evol}}$ is then simply defined according to \citet{Peters1964} as
\begin{equation}
    a_f = (a_i - 4\beta t_\text{evol})^{1/4},
\end{equation}
where $a_i$ is the DWD separation at formation, and 
\begin{equation}
    \beta = \frac{64G^3}{5c^5} M_1M_2(M_1+M_2)
\end{equation}
is constant throughout DWD evolution \citep{Peters1964}.

@@

\begin{equation}
    a_{\text{RLO,}\ell} = R_{\ell} \frac{0.6 q_{\ell}^{2/3} + \ln{(1+q_{\ell}^{1/3})}}{0.49 q_{\ell}^{2/3}}
\end{equation}
where $R_{\ell}$ is the radius of the lower mass WD and $q_{\ell} = M_{\ell}/M_{h}$ is the ratio of the lower- to higher-mass WD components. \citep{Eggleton1983}. 
We define the radius of a WD following \citet{Tout1997, Hurley2000} as
\begin{equation}
    R_{\rm{WD}} = \max\Bigg(R_\text{NS}, 0.0115\sqrt{\left(\frac{M_\text{Ch}}{M}\right)^{2/3}-\left(\frac{M}{M_\text{Ch}}\right)^{2/3}} \Bigg) 
\end{equation}
\noindent where $R_\text{NS} = 1.4\cdot 10^{-5}$ R$_\odot$ is the radius of a neutron star, $M_{\text{Ch}}=1.44$ M$_\odot$ is the Chandrasekhar limit for the mass of a stable WD, and $M$ is the mass of the DWD in solar masses.

For the non-discarded systems, we log the present-day separations from which the present-day orbital frequency $f_\text{orb}$ can be found using Kepler's third law. The GW frequency is then $f_\text{GW} = 2f_\text{orb}$.
